Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants

### 2016-10-15 paper/paper.tex
nudges

\begin{abstract}
%
Modern control is implemented using digital microcontrollers, which are
designed to work in an embedded environment, within a dynamical plant that
represents physical components.
%
We present a new algorithm based on counterexample guided inductive
synthesis that automates the design of digital controllers that are
correct by construction.  The synthesis result is sound with respect to the
complete range of approximations, including time discretization,
quantization effects, and finite-precision arithmetic and their related
rounding errors.
%
We have implemented our new algorithm in a tool called \tool, and are able
to automatically generate stable controllers for a set of intricate plant
models taken from the literature within minutes.
%
\end{abstract}

@@

Beyond classical a-posteriori validation procedures in digital control, 
there has been plenty of previous work aiming at \emph{verifying} a given
controller designed by the engineer.  Recent work can show the stability of 
digital controllers with considerations to implementation aspects, i.e.,
fixed-point arithmetic and the word length used~\cite{Bessa16}.  This work
exploits advances in bit-accurate verification of C programs to obtain a
verifier for software-implemented digital control.

By contrast, we leverage a very recent step-change in the scalability of
\emph{program synthesis}.  Program synthesis engines use a program
specification as the starting point, and subsequently generate a sequence of
candidate programs.  The candidate programs are iteratively refined to match
the specification more closely.  Program synthesizers that implement
Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} are now
able to generate programs for highly non-trivial specifications with a very
high degree of automation.  Modern CEGIS engines implement combinations of
automated testing, genetic algorithms and SMT-based automated
reasoning~\cite{DBLP:conf/lpar/DavidKL15, DBLP:conf/cav/0001A14}.

@@

Our work addresses particularly challenging aspects of the control synthesis
problem.  We perform digital control synthesis over a hybrid model, where
the plant exhibits continuous behavior whereas the controller operates in 
discrete time and over a quantized domain.  
Inspired by a classical approach \cite{astrom1997computer}, 
we translate the problem into a single digital domain.  
Given our primary interest in stability despite digitization effects, we have modeled
a digital equivalent of the plant by evaluating the effects of the
quantizers (A/D and D/A converters) as time discretization elements with
quantization noise.  The resulting closed-loop system is a program loop that
operates on bit-vectors encoded using fixed-point arithmetic with finite-word
length (FWL). The deterministic effects of the FWL and
quantization errors are incorporated into the model as an additive
uncertainty, which is taken into account during the CEGIS-based synthesis of
the control software for the plant.

@@

\item We design and implement an automatic approach in a tool called \tool  
  for generating {\em corect-by-construction} digital controllers that is based on
  state-of-the-art inductive synthesis techniques.  Our application of
  program synthesis is non-trivial and requires addressing challenges
  specific to embedded control systems, such as the effects of quantizers and
  FWL. In particular, we have found that a two-stage verification engine
  that continuously refines the precision of the fixed-point
  representation of the plant is key to the performance of our approach yielding,
  in most cases, a speed-up of at least two orders of magnitude over a
  conventional one-stage verification engine.

@@

%% \item We exploit a CEGIS-based approach that uses inductive synthesis
%% together with a bit-precise formal verification algorithm for synthesising
%% correct-by-design robust digital controllers for non-fragile closed-loop
%% stability.

%% \item Experimental results with our synthesiser show that the
%% combination of interval arithmetic to represent the controller coefficients
%% and a two-stage verification approach to continuously refine precision
%% enables us to compute a parametric, sound solution for $14$ out of $15$
%% digital stabilizing controllers for a set of continuous and discrete plants
%% with parametric uncertainties.

@@

%\textcolor{green}{DC: I believe Alessandro wanted to be more specific about the nature of the modelling error which is why I was trying to explain it. In this case I believe there are also two separate sources for $\Delta{G}_p(z)$. The first is $\Delta{G}_p(s)$ (which is basically a known tolerance that may come from G(t)), whilst the second is the error introduced by the calculations in the $\mathcal{Z}$ transform. Alessandro, do you want to address this second term (possibly as explained below) or just leave the blue section as is?}

%Since $G(z)$ is an approximation of $G(s)$ we introduce an error term caused during its calculation as $\Delta{G}_P$
%
%%
%\begin{align}
%\small
%\label{parametric_plant_tf}
%\hat{G}(z)&=\frac{(b_{0}+\Delta_p b_{0}) +...+(b_{M_G}+\Delta_p b_{M_G})z^{-M_G}}{(a_{0}+\Delta_p a_{0})+...+(a_{N_G}+\Delta_p a_{N_G})z^{-N_G}} \nonumber \\
%&=G(z)+\Delta{G}_P(z).
%\end{align}
%%
%Note that $\Delta{G}_P$ is undefined and upperbound by a user defined
%value such that $\hat{G}(z)$ corresponds to the set of
%possible polynomial forms of $G(z, T)$ that overapproximate $G(s)$.
%This calculation is performed outside of the scope of this work since it depends on the nature of the tool used to obtain the $\mathcal{Z}$ form. The error comes from the bounded horizon and mathematical precision used to calculate the $\mathcal{Z}$ transform, which will be tool and user dependent.
%\red{[The previous statement is confusing and I must disagree here: I want hard error bounds on a specific value of T.]}
%\green{[You are correct. It was absolutely wrong. Hopefully this explanation is clear. We use MATLAB, but still don't know the details of their implementation, so I don't have hard data on how this error is calculated.}

@@

%\begin{mylemma}
%\label{lemma:variance}
%%
%In particular, Widrow \cite{widrow1956} models the quantization noise as a
%gaussian model in which the second moment (variance) of the probability
%density function (PDF) of the quantization noise is given by
%%
%\begin{equation}
%\label{eq:variancelemma}
%\sigma_{Q}^{2}=\frac{q^{2}}{12},
%\end{equation}
%where $q$ is the step size of quantiser. (Proof is on \cite{widrow1956})
%\end{mylemma}
%
%\begin{mylemma}
%\label{lemma:boundednoise}
%%
%The power spectral density of the quantization noise (supposed white)
%$S_{\nu}(z)$ is bounded and may be computed through the z-transform of
%autocorrelation function of $\nu(k)$, thus:
%%
%\begin{equation}
%S_{\nu}(z)= \sigma_{Q}^{2}=\frac{q^{2}}{12}.
%\end{equation}
%\end{mylemma}
%
%\begin{proof}
%%
%Considering $\nu(k)$ a random process, where $\nu(k)$ is independent from
%$\nu(k+i)$ for every integer $k\neq 0$.  The autocorrelation function of
%$\nu(k)$, {\it i.e.} $R_{\nu\nu}(\tau)$ may be computed as follows:
%%
%\begin{equation}
%R_{\nu\nu}(\tau)=E[\nu(t+\tau)\nu(t)]=\sigma_{Q}^{2}\cdot \delta(\tau),
%\end{equation}
%%
%where $\delta(\tau)$ is the Dirac delta function at delay time $\tau$.
%
%By definition~\cite{poularikas2000transforms}, the power spectral density of
%$\nu(k)$ is the z-transform of $R_{\nu\nu}(\tau)$, thus
%%
%\begin{equation}
%S_{\nu}(z)=  \sum\limits_{k=-\infty}^{\infty}{R_{\nu\nu}(k)z^{-k}} = \sigma_{Q}^{2},
%\end{equation}
%considering the Lemma~\ref{lemma:variance}:
%\begin{equation}
%S_{\nu}(z)=  \frac{q^{2}}{12}.
%\end{equation}
%\end{proof}
%
%Finally, the quantization noise spectrum may be considered a energy-bounded
%(Lemma~\ref{lemma:boundednoise}), and consequently the Parseval's theorem
%ensures that the quantization noise is also energy-bounded in time-domain. 
%Thus $\hat{Y}(z)$ will be bounded and the closed-loop system stable if the
%conditions in Lemma~\ref{lemma:eq_int_stab} are followed.

@@

% \blue{Initially, the controller denominator and numerator  ($D_{C}(z)$ and $N_{C}(z)$) are started with zero coefficients, as in previous example. Thus,  in the first {\sc verify} phase a counterexample is presented, that is added to {\sc inputs}, whis is usede in the first  {\sc synthesise} phase. The resulting controller is:
% }
% $$
% {
% \blue{
% C(z)=\frac{0.77516z^2{+}-0.71497z{+}0.03729}{0.25024z^2{-}0.01799z{+}0.00097}
% }}
% $$

@@

% \blue{
% This candidate solution is validated as the final solution by the next 
% {\sc verify} phase.
% }
%
\begin{figure}[htb]
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd0.pdf}
        \caption{Original controller~\cite{DBLP:conf/hybrid/WangGRJF16}.}
        \label{fig:original}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd2.pdf}
        \caption{Controller synthesized by \tool.}
        \label{fig:cegiscontroller}
    \end{subfigure}
    \caption{Bode diagram for original~\cite{DBLP:conf/hybrid/WangGRJF16} and synthesized closed-loop systems.}\label{fig:bode}
\end{figure}
%
Figure~\ref{fig:step} shows the step responses of the closed-loop system 
with the original controller represented by Eq.~\eqref{Eq:running-example-controller} (Figure~\ref{fig:step0}), 
the first (Figure~\ref{fig:step1}) and final (Figure~\ref{fig:step2}) candidate solutions provided by \tool. 
The step response in Figure~\ref{fig:step0} confirms the stability loss if we consider FWL effects. 
However, Figure~\ref{fig:step1} shows that the first candidate controller is able to stabilize 
the closed-loop system without uncertainties, but it is rejected during the {\sc precision} phase by 
\tool since this solution is not considered to be sound. Finally, Figure~\ref{fig:step2} 
shows a stable behavior for the final (sound) solution, which presents a lower settling time.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step0.pdf}
        \caption{Orignal controller}
        \label{fig:step0}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step1.pdf}
        \caption{First solution by \tool}
        \label{fig:step1}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step2.pdf}
        \caption{Final solution by \tool}
        \label{fig:step2}
    \end{subfigure}
    \caption{Step responses for original~\cite{DBLP:conf/hybrid/WangGRJF16}
             closed-loop system with FWL effects and for each
             {\sc synthesize} iteration of \tool}\label{fig:step}
\end{figure*}

@@

Researchers in formal verification of control systems have studied various
effects of discretizing the dynamics, including delayed
response~\cite{Duggirala2015}, and Finite Word Length (FWL) constraints with
the goal to either verify (e.g.~\cite{daes20161}) or optimize
(e.g.~\cite{oudjida2014design}) implementations.  There are in fact two
different problems arising from FWL effects.  The first corresponds to the
error in the dynamics caused by the inability of the hardware to represent
the exact dynamics of the system whilst the second relates to rounding
errors during computation.  In~\cite{fialho1994stability} a stability
measure based on the error of the digital dynamics ensures that the FWL
errors do not make the digital system unstable.  A more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  Most studies following this verification research field
focus on finding a correct implementation of a known desired controller,
looking for optimal parameter representations using FWL, but ignoring the
effects of rounding errors due to issues of mathematical tractability.  The
analyses in~\cite{DBLP:conf/hybrid/WangGRJF16, DBLP:conf/hybrid/RouxJG15}
rely on an invariant computation on the discrete system dynamics using
Semi-Definite Programming (SDP).  Whilst the former uses BIBO properties to
determine stability, the latter uses Lyapunov-based quadratic invariants. 
In both cases, the SDP solver uses floating-point arithmetic and soundness
is checked by upper bounding the error.  An alternative approach is taken
by~\cite{park2016scalable}, where the verification of existing code is
performed against a known model by extracting an LTI model of the code
through symbolic execution.  In order to account for rounding errors, an
upper bound is introduced in the verification phase.  If the error of the
implementation is lower than this tolerance level, then the verification is
successful.

@@

Given the current knowledge in control system's synthesis, there is no other
existing study on the automatic synthesis of fixed-point digital controllers
considering FWL effects.  Parameter synthesis tools such
as~\cite{cimatti2013parameter} use bounded model checking and fixed-point
computations to find parameters based on user defined specifications, but
they often operate in the continuous domain and are ill-suited for robust
analysis since they have no direct means of evaluating robustness.  Other
tools such as~\cite{economakos2016automated} are directly aimed at robust
stability problems, but they fail to explore the effects of FWL in their
implementation.  In order to provide a correct by design digital controller,
\cite{alur2016compositional} requires a user-defined finite state
abstraction to synthesize a digital controller based on high level
specifications.  Whilst this approach directly overcomes the challenges
presented by the FWL problem, it still requires error-prone user
intervention.  A different solution that looks at the other end of the scale
is an approach that synthesizes word lengths for known problems as developed
in~\cite{jha2013swati}; however, this provides neither an optimal word
synthesis nor a comprehensive solution for the problem.  It is from this
vacuum that we seek to explore a CEGIS approach to FWL controller synthesis.

@@

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high-level specifications.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesize invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesize.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{DBLP:conf/lpar/DavidKL15} and the latter from~\cite{daes20161}
whilst enhancing the procedure by evaluating the quantization effects of
the Hardware interfaces (ADC/DAC) to obtain an accurate discrete-time FWL
representation of the continuous dynamics.

>>>>

\begin{abstract}
%
Modern control is implemented with digital microcontrollers, which are
designed to work in an embedded environment, within a dynamical plant that
represents physical components.
%
We present a new algorithm based on counter\-example guided inductive
synthesis that automates the design of digital controllers that are
correct by construction.  The synthesis result is sound with respect to the
complete range of approximations, including time discretization,
quantization effects, and finite-precision arithmetic and their related
rounding errors.
%
We have implemented our new algorithm in a tool called \tool, and are able
to automatically generate stable controllers for a set of intricate plant
models taken from the literature within minutes.
%
\end{abstract}

@@

Beyond classical a-posteriori validation procedures in digital control,
there has been plenty of previous work aiming at \emph{verifying} a given
controller designed by the engineer.  Recent work can show the stability of
digital controllers with consideration of implementation aspects, i.e.,
fixed-point arithmetic and the word length used~\cite{Bessa16}.  The authors
of~\cite{Bessa16} exploit advances in bit-accurate verification of C
programs to obtain a verifier for software-implemented digital control.

By contrast, we leverage a very recent step-change in the scalability of
\emph{program synthesis}.  Program synthesis engines use a program
specification as the starting point, and subsequently generate a sequence of
candidate programs.  The candidate programs are iteratively refined to match
the specification more closely.  Program synthesizers that implement
Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} are now
able to generate programs for highly non-trivial specifications with a very
high degree of automation.  Modern CEGIS engines combine automated testing,
genetic algorithms and SMT-based automated
reasoning~\cite{DBLP:conf/lpar/DavidKL15, DBLP:conf/cav/0001A14}.

@@

Our work addresses particularly challenging aspects of the control synthesis
problem.  We perform digital control synthesis over a hybrid model, where
the plant exhibits continuous behavior whereas the controller operates in
discrete time and over a quantized domain.  Inspired by a classical approach
\cite{astrom1997computer}, we translate the problem into a single digital
domain.  Given our primary interest in stability despite digitization
effects, we have modeled a digital equivalent of the plant by evaluating the
effects of the quantizers (A/D and D/A converters) as time discretization
elements with quantization noise.  The resulting closed-loop system is a
program loop that operates on bit-vectors encoded using fixed-point
arithmetic with finite word length (FWL).  The deterministic effects of the
FWL and quantization errors are incorporated into the model as an additive
uncertainty, which is taken into account during the CEGIS-based synthesis of
the control software for the plant.

@@

\item We design and implement an automatic approach in a tool called \tool  
  for generating {\em correct-by-construction} digital controllers that is
  based on state-of-the-art inductive synthesis techniques.  Our application
  of program synthesis is non-trivial and requires addressing challenges
  specific to embedded control systems, such as the effects of quantizers
  and FWL.  In particular, we have found that a two-stage verification
  engine that continuously refines the precision of the fixed-point
  representation of the plant is key to the performance of our approach
  yielding, in most cases, a speed-up of at least two orders of magnitude
  over a conventional one-stage verification engine.

@@

%% \item We exploit a CEGIS-based approach that uses inductive synthesis
%% together with a bit-precise formal verification algorithm for synthesizing
%% correct-by-design robust digital controllers for non-fragile closed-loop
%% stability.

%% \item Experimental results with our synthesizer show that the
%% combination of interval arithmetic to represent the controller coefficients
%% and a two-stage verification approach to continuously refine precision
%% enables us to compute a parametric, sound solution for $14$ out of $15$
%% digital stabilizing controllers for a set of continuous and discrete plants
%% with parametric uncertainties.

@@

%\textcolor{green}{DC: I believe Alessandro wanted to be more specific about the nature of the modelling error which is why I was trying to explain it. In this case I believe there are also two separate sources for $\Delta{G}_p(z)$. The first is $\Delta{G}_p(s)$ (which is basically a known tolerance that may come from G(t)), while the second is the error introduced by the calculations in the $\mathcal{Z}$ transform. Alessandro, do you want to address this second term (possibly as explained below) or just leave the blue section as is?}

%Since $G(z)$ is an approximation of $G(s)$ we introduce an error term caused during its calculation as $\Delta{G}_P$
%
%%
%\begin{align}
%\small
%\label{parametric_plant_tf}
%\hat{G}(z)&=\frac{(b_{0}+\Delta_p b_{0}) +...+(b_{M_G}+\Delta_p b_{M_G})z^{-M_G}}{(a_{0}+\Delta_p a_{0})+...+(a_{N_G}+\Delta_p a_{N_G})z^{-N_G}} \nonumber \\
%&=G(z)+\Delta{G}_P(z).
%\end{align}
%%
%Note that $\Delta{G}_P$ is undefined and upper bound by a user defined
%value such that $\hat{G}(z)$ corresponds to the set of
%possible polynomial forms of $G(z, T)$ that overapproximate $G(s)$.
%This calculation is performed outside of the scope of this work since it depends on the nature of the tool used to obtain the $\mathcal{Z}$ form. The error comes from the bounded horizon and mathematical precision used to calculate the $\mathcal{Z}$ transform, which will be tool and user dependent.
%\red{[The previous statement is confusing and I must disagree here: I want hard error bounds on a specific value of T.]}
%\green{[You are correct. It was absolutely wrong. Hopefully this explanation is clear. We use MATLAB, but still don't know the details of their implementation, so I don't have hard data on how this error is calculated.}

@@

%\begin{mylemma}
%\label{lemma:variance}
%%
%In particular, Widrow \cite{widrow1956} models the quantization noise as a
%Gaussian model in which the second moment (variance) of the probability
%density function (PDF) of the quantization noise is given by
%%
%\begin{equation}
%\label{eq:variancelemma}
%\sigma_{Q}^{2}=\frac{q^{2}}{12},
%\end{equation}
%where $q$ is the step size of quantiser. (Proof is on \cite{widrow1956})
%\end{mylemma}
%
%\begin{mylemma}
%\label{lemma:boundednoise}
%%
%The power spectral density of the quantization noise (supposed white)
%$S_{\nu}(z)$ is bounded and may be computed through the z-transform of
%autocorrelation function of $\nu(k)$, thus:
%%
%\begin{equation}
%S_{\nu}(z)= \sigma_{Q}^{2}=\frac{q^{2}}{12}.
%\end{equation}
%\end{mylemma}
%
%\begin{proof}
%%
%Considering $\nu(k)$ a random process, where $\nu(k)$ is independent from
%$\nu(k+i)$ for every integer $k\neq 0$.  The autocorrelation function of
%$\nu(k)$, {\it i.e.} $R_{\nu\nu}(\tau)$ may be computed as follows:
%%
%\begin{equation}
%R_{\nu\nu}(\tau)=E[\nu(t+\tau)\nu(t)]=\sigma_{Q}^{2}\cdot \delta(\tau),
%\end{equation}
%%
%where $\delta(\tau)$ is the Dirac delta function at delay time $\tau$.
%
%By definition~\cite{poularikas2000transforms}, the power spectral density of
%$\nu(k)$ is the z-transform of $R_{\nu\nu}(\tau)$, thus
%%
%\begin{equation}
%S_{\nu}(z)=  \sum\limits_{k=-\infty}^{\infty}{R_{\nu\nu}(k)z^{-k}} = \sigma_{Q}^{2},
%\end{equation}
%considering the Lemma~\ref{lemma:variance}:
%\begin{equation}
%S_{\nu}(z)=  \frac{q^{2}}{12}.
%\end{equation}
%\end{proof}
%
%Finally, the quantization noise spectrum may be considered a energy-bounded
%(Lemma~\ref{lemma:boundednoise}), and consequently the Parseval's theorem
%ensures that the quantization noise is also energy-bounded in time-domain. 
%Thus $\hat{Y}(z)$ will be bounded and the closed-loop system stable if the
%conditions in Lemma~\ref{lemma:eq_int_stab} are followed.

@@

% \blue{Initially, the controller denominator and numerator  ($D_{C}(z)$ and $N_{C}(z)$) are started with zero coefficients, as in previous example. Thus,  in the first {\sc verify} phase a counterexample is presented, that is added to {\sc inputs}, whis is used in the first  {\sc synthesise} phase. The resulting controller is:
% }
% $$
% {
% \blue{
% C(z)=\frac{0.77516z^2{+}-0.71497z{+}0.03729}{0.25024z^2{-}0.01799z{+}0.00097}
% }}
% $$

@@

% \blue{
% This candidate solution is validated as the final solution by the next 
% {\sc verify} phase.
% }
%
\begin{figure}[htb]
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd0.pdf}
        \caption{Original controller~\cite{DBLP:conf/hybrid/WangGRJF16}.}
        \label{fig:original}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd2.pdf}
        \caption{Controller synthesized by \tool.}
        \label{fig:cegiscontroller}
    \end{subfigure}
    \caption{Bode diagram for original~\cite{DBLP:conf/hybrid/WangGRJF16} and synthesized closed-loop systems.}\label{fig:bode}
\end{figure}
%
Figure~\ref{fig:step} shows the step responses of the closed-loop system 
with the original controller represented by Eq.~\eqref{Eq:running-example-controller} (Figure~\ref{fig:step0}), 
the first (Figure~\ref{fig:step1}) and final (Figure~\ref{fig:step2}) candidate solutions provided by \tool. 
The step response in Figure~\ref{fig:step0} confirms the stability loss if we consider FWL effects. 
However, Figure~\ref{fig:step1} shows that the first candidate controller is able to stabilize 
the closed-loop system without uncertainties, but it is rejected during the {\sc precision} phase by 
\tool since this solution is not considered to be sound. Finally, Figure~\ref{fig:step2} 
shows a stable behavior for the final (sound) solution, which presents a lower settling time.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step0.pdf}
        \caption{Original controller}
        \label{fig:step0}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step1.pdf}
        \caption{First solution by \tool}
        \label{fig:step1}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step2.pdf}
        \caption{Final solution by \tool}
        \label{fig:step2}
    \end{subfigure}
    \caption{Step responses for original~\cite{DBLP:conf/hybrid/WangGRJF16}
             closed-loop system with FWL effects and for each
             {\sc synthesize} iteration of \tool}\label{fig:step}
\end{figure*}

@@

Researchers in formal verification of control systems have studied various
effects of discretizing the dynamics, including delayed
response~\cite{Duggirala2015}, and Finite Word Length (FWL) constraints with
the goal to either verify (e.g.~\cite{daes20161}) or optimize
(e.g.~\cite{oudjida2014design}) implementations.  There are in fact two
different problems arising from FWL effects.  The first corresponds to the
error in the dynamics caused by the inability of the hardware to represent
the exact dynamics of the system while the second relates to rounding
errors during computation.  In~\cite{fialho1994stability} a stability
measure based on the error of the digital dynamics ensures that the FWL
errors do not make the digital system unstable.  A more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  Most studies following this verification research field
focus on finding a correct implementation of a known desired controller,
looking for optimal parameter representations using FWL, but ignoring the
effects of rounding errors due to issues of mathematical tractability.  The
analyses in~\cite{DBLP:conf/hybrid/WangGRJF16, DBLP:conf/hybrid/RouxJG15}
rely on an invariant computation on the discrete system dynamics using
Semi-Definite Programming (SDP).  While the former uses BIBO properties to
determine stability, the latter uses Lyapunov-based quadratic invariants. 
In both cases, the SDP solver uses floating-point arithmetic and soundness
is checked by upper bounding the error.  An alternative approach is taken
by~\cite{park2016scalable}, where the verification of existing code is
performed against a known model by extracting an LTI model of the code
through symbolic execution.  In order to account for rounding errors, an
upper bound is introduced in the verification phase.  If the error of the
implementation is lower than this tolerance level, then the verification is
successful.

@@

Given the current knowledge in control system's synthesis, there is no other
existing study on the automatic synthesis of fixed-point digital controllers
considering FWL effects.  Parameter synthesis tools such
as~\cite{cimatti2013parameter} use bounded model checking and fixed-point
computations to find parameters based on user defined specifications, but
they often operate in the continuous domain and are ill-suited for robust
analysis since they have no direct means of evaluating robustness.  Other
tools such as~\cite{economakos2016automated} are directly aimed at robust
stability problems, but they fail to explore the effects of FWL in their
implementation.  In order to provide a correct by design digital controller,
\cite{alur2016compositional} requires a user-defined finite state
abstraction to synthesize a digital controller based on high level
specifications.  While this approach directly overcomes the challenges
presented by the FWL problem, it still requires error-prone user
intervention.  A different solution that looks at the other end of the scale
is an approach that synthesizes word lengths for known problems as developed
in~\cite{jha2013swati}; however, this provides neither an optimal word
synthesis nor a comprehensive solution for the problem.  It is from this
vacuum that we seek to explore a CEGIS approach to FWL controller synthesis.

@@

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high-level specifications.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesize invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesize.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{DBLP:conf/lpar/DavidKL15} and the latter from~\cite{daes20161}
while enhancing the procedure by evaluating the quantization effects of
the Hardware interfaces (ADC/DAC) to obtain an accurate discrete-time FWL
representation of the continuous dynamics.
